Do educational institutions explain SES gaps?: an interntational comparison of the SES achievement gap from 2000-2015

### 2017-09-12 paper/paper.tex
Finalized first draft of paper - ran everything: good

\usepackage{geometry}
 \geometry{
 a4paper,
 total={170mm,257mm},
 left=20mm,
 top=20mm,
 }
\usepackage{pdflscape}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[round, colon]{natbib}
\usepackage[colorlinks]{hyperref}
\usepackage{booktabs}
\AtBeginDocument{%
  \hypersetup{
    citecolor=blue,
    linkcolor=blue,   
    urlcolor=blue}}

\title{The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015}
\author{Jorge Cimentada}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\setlength{\parindent}{2em}
\setlength{\parskip}{1em}
\showboxdepth=5
\showboxbreadth=5

@@

On the other hand, we do have some countries which show a decrease in the SES achievement gap. Poland decreased by about -0.4 and Denmark by -0.2. However, the most notable cases are the United States and Germany. These two countries show a very high level of dispersion in the year 2000 with SES gaps of over 2 standard deviations. But in the 15-year time trend both countries reduced the gaps by -0.6 and -0.9 respectively. Their distinctively big gaps in 2000 also show up in the work of \citet{micklewright}. This corroborates the findings of \citet{reardon_portilla}, which found a decreasing gap for kindergarteners. This decline is evident also for 15 year olds, suggesting it might be more of an institutional change rather than a specific grade-level policy. Do note that the trendline is for both mathematics and literacy. If it were only for mathematics, then the trendline would probably be even more extreme. Despite this, pooling both subjects gives even more strength to the analysis given that the results stand for both subjects together \footnote{Same graph is availabe for reading upon request, where we find pretty much the same results if not even more steep/downward slopes.}. We also plot the 80/20 and 70/30 SES gaps and find that very similar patterns as in the 90/10 SES gap (see figure 10 in the appendix).

@@

We find that the initial gap for the U.S in 2000 is 2.17 varying between 2.13 and 2.21. This gap is much higher than found in previous studies, but again, we don't know how much previous estimates vary and nobody has really studied the 15-year old achievement gap precisely. All country-level studies pool many studies together and control for age, removing the age-specific gap effect.

Figure 3 below takes a more direct approach and looks at the percentage change from the first and last timepoint available. Each data point has been computed together with it's 50\% and 95\% uncertainty interval \footnote{Each of these uncertainty intervals were computed using a 500-replicate bootstrap}.

@@

The U.S shows a different pattern from all other countries. Both the 90th/50th/10th gaps seem to contribute rather similarly to the overall 90th/10th gap (although 50th/10th is marginally higher) but more interestingly, the 90th/50th gap has stagnated in the last 15 years. The top/middle groups have a relatively big gap and it hasn't changed at all. On the other hand, the steep decrease we saw for the U.S is coming exclusively from the 50th/10th gap shrinking. The reason, nonetheless, might be different from the German case, where it's probable that the bottom is catching up. Here, as in the 90th/10th gap, the middle might be 'dumbing' down closer to the low SES group. The big takeaway from the U.S is that the decrease of inequality is coming from the shrinking of the 50/10 gap.

@@

To sum up, we find no evidence in support of our first hypothesis. We uncover that the tracking setup of a country doesn't imminently mean a certain degree of equality/inequality. Countries where the top SES group is increasing and the bottom is decreasing (top-left panel) have dramatically different tracking setups. Austria is an early tracker with several tracks, while Sweden is a late tracker with only 1 track. On the opossite panel (bottom-right), where we find countries that the bottom group is catching up while the top is coming down, we find similar results. Denmark and Germany are opposite poles in terms of tracking, while the U.S is distinctive in that it has little formal curricular stratification. We see no definitive evidence of top-to-bottom equality in all of these countries (except for Poland). We find that regardless of the different tracking setups, we see that there's a tradeoff: either the bottom comes up and the top comes down or the top goes even further up and the bottom further down.

@@

Moving on, we now discuss if there's a tradeoff between an educational system being efficient, i.e raising the overall performance of the country and whether that challenge compromises equality at the same time. For example, in order to improve the average score in, let's say, mathematics, tracking can improve the average score by making children in the higher trackers perform better, raising the overall performance. This, however, can come at a price. If tracking actually benefits the top but doesn't increase the performance of the bottom tracks then we see a net increase of inequality accompanied with an increase in the overall performance. Given that we have the data for the average performance for each country and the achievement gap, we can investigate this relationship in detail. We first being by plotting all countries pooled for each year in figure 8.

@@

For the past 15 years we see a very strong correlation suggesting that countries that perform very well in PISA have actually very low achievement gaps (or the other way around) \footnote{We graph all countries-years pooled as well and the relationship is identical}. The correlations are between -0.73 and -0.88. This confirms the theoretical and empirical frameworks suggested by \citet{werfhorst_mijs}. However, what they didn't really go in depth was whether this corelation is strongest for some countries over time. However, we should not get trapped into an ecological fallacy and assume this relationship is identical within each country. With the data available we can look at the evolution separately for many countries. We plot it up next in figure 9.

@@

In the final model, we regress the inequality of a country on a dummy variable where 1 equals only 1 track vs more than 1 track, whether the age of first selection is 15 or more vs below 15, a dummy stating whether the percentage of tracked curriculum is above 0 (that is, any tracking), and the standardized vocational index \footnote{We create dummies to avoid multicollinearity. The final model has a maximum VIF of 2.2.}. We are reluctant to include more variable in the model, first to keep it more parsimonious, and secondly to prevent overfitting due to our sample size. We present the results in the table below.

@@

Given that the classical p values are inexistent in bayesian modelling, we present uncertainty intervals next to each coefficient. The first model only includes the 1 track dummy. This variable shows that countries that have one track, have on average, half a standard deviation less inequality than other countries. This covariate alone explains about 23\% of the variance of the evolution of the 90/10 achievement gap. The second model now includes the dummy of whether the country has an age selection of 15 or more and we can see that countries with late selection into tracking have about -.38 less standard deviations in inequality. Also note that the number of tracks dummy was now reduced, suggesting these two variables are explaining similar things. This is expected given that late tracking is usually associated with fewer tracks, such as in the Scandinavian countries. The third model now includes whether the country has any degree of tracking, meaning whether the country has above 0\% of the curriculum tracked. The coding of the variable comes from visual inspection of the distribution, where some countries, such as the United States and Canada, have untracked curriculums. For this last variable we see no relationship at all, but after including the vocational index, the model is completely reversed with countries with tracked curriculum having about 0.6 more standard deviations in inequality that countries with no tracking.

We also note that vocational enrollment is associated with less inequality. Not surprisingly, the age of selection coefficient, also increased due to the correlation between age of selection and vocational enrollment. Countries with high vocational enrollment are also those with early age selection, such as Germany and Austria. These two variables should be inspected further, perhaps with an interaction, given that both coefficients are sizable and very correlated. This final model explains over one third of the variance in the evolution of the 90/10 SES gap, with an R square of 40\%. The uncertainty of each of the covariates varies a lot with the percentage of curriculum tracked dummy ranging all the way from 0.9 to 0.3. Nonetheless, we see that the other coefficients are very strong and have reasonable intervals. It is very important to consider the uncertainty of our estimates, to gain confidence in our model fit. So far, this model seems to fit the data quite reasonably.

All of these countries have a different tracking structure, with some setups being more egalitarian than others. Using the previous model we make a simulation and predict the level of the 90/10 achievement gap for all possible tracking setups for all countries. Below we plot all these combinations. The X axis has the number of tracks that a country has, the Y axis has the predicted 90/10 achievement gap and below each country name we have whether the country has an age selection of 15 or more or below 15. In addition, to confirm how accurate the model is, we also plot the actual level of inequality in the actual tracking setup of the country. Let's take Germany for example.

@@

These graph reveals some interesting patterns. Generally speaking, we find that the simulation predicts that virtually all countries which are not in this 'ideal' tracking, that would switch to the 'ideal' setup, would experience a reduction of the achievement gap. Yet we also find the opposite result. Countries which are in the 'ideal' setup that would change to the 'worst' setup would see a widening of the achievement gap. On average, we find that if all countries switched to the 'ideal' setup, we would see an average reduction of 20\%. This is a very important figure considering that \citet{reardon2011} found that the US gap increased by about 40-50\% in 30 years. Had all countries in the 'ideal' tracking switched to the 'worst' tracking, we would see an average widening of nearly 72\%. For those countries which are not in the 'ideal' tracking setup, we compute below the percentage change from switching to the ideal tracking.

@@

A more summarized empirical strategy is to actually standardize all the tracking indicator into one index, something that \citet{bol2013} already do quite nicely. Using this standardized tracking variable and the same bayesian multilevel model, we now model the tracking and vocational index as before, but add an interaction between two, given that in the previous model we recorded a strong shift in the model when tracking-related variables and the vocational model where both included in the same model. Below are the results for this model.

@@

We confirm the results from the previous models, where more tracking is associated with an achievement gap of about 0.4 standard deviations wider and more vocational enrollment is associated with a reduction of about 0.2 standard deviations, half the slope of tracking. Meaning that the tracking effect might be offsetting the effect of the vocational index given that it's half the size of the slope. Both these coefficients are actually significantly different from each other, so it is possible. Before we explore it visually, note that the interaction between the two is very reliable and significant, considering that small sample size. Below we graph the interaction for different quantiles of the vocational index.

@@

{\centering \includegraphics[width=5.5in,height=5in]{figure/interaction_plot-1}

@@

The results are very interesting. We see that for lower levels of tracking (meaning little to no tracking) theres a very low achievement gap, regardless of whether the country has very high or very low vocational enrollment. However, once tracking enters the picture, vocational enrollment can be a strong equalizer. We see that the bottom quantile of vocational enrollment has almost twice the achievement gap as the top quantile for vocational enrollment. Moreover, our earlier interpretation (that tracking is more important than vocational enrollment for reducing the achievement gap) still holds and is very evident here. Despite the equalizing power of vocational enrollment, a high level of tracking with the highest level of vocational enrollment still leaves a country with an achievement gap of over 1.5 standard deviations. However, a country with low levels of tracking and \textit{any} level of vocational enrollment is nearly 1.3 standard deviations at best. A key finding from these results is that tracking is a reasonable explanation of the 90/10 achievement gap, and we should consider it as a very important component in the reduction of the achievement gap. However, these results highlight that it should not be studied in isolation from other features such as the degree of vocational enrollment. We see a very sizable change in the achievement gap for a given country with the same level of tracking and high and low levels of vocational enrollment.

@@

To test whether these results hold for other samples, we compute the same models as before but also for the 80/20 and 70/30 SES gap. Up next we plot these coefficients for the three SES gaps.

@@

Instead of estimating the evolution over time, we calculate the cumulative change over time for each country, summarizing the 6 year-time span of achievement gaps into one cumulative inequality gap. This approach dramatically reduces the total sample size to 34, the number of total countries available \footnote{We've been presenting a selected number of countries given the lack of space, but for every model/estimation we include all countries available.}. It is for this reason that we only include at most 2 covariates in the model, the tracking and vocational index, which both summarizes the previous models and keeps it parsimonious.

@@

@@

>>>>

\usepackage{geometry}
 \geometry{
 a4paper,
 total={170mm,257mm},
 left=20mm,
 top=20mm,
 }
\usepackage{pdflscape}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{titling}
\usepackage[round, colon]{natbib}
\usepackage[colorlinks]{hyperref}
\usepackage{booktabs}
\AtBeginDocument{%
  \hypersetup{
    citecolor=blue,
    linkcolor=blue,   
    urlcolor=blue}}

\title{Do educational institutions explain SES gaps?: an interntational comparison of the SES achievement gap from 2000-2015}
\author{Jorge Cimentada}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\setlength{\parindent}{2em}
\setlength{\parskip}{1em}
\showboxdepth=5
\showboxbreadth=5

@@

On the other hand, we do have some countries which show a decrease in the SES achievement gap. Poland decreased by about -0.4 and Denmark by -0.2. However, the most notable cases are the United States and Germany. These two countries show a very high level of dispersion in the year 2000 with SES gaps of over 2 standard deviations. But in the 15-year time trend both countries reduced the gaps by -0.6 and -0.9 respectively. Their distinctively big gaps in 2000 also show up in the work of \citet{micklewright}. This corroborates the findings of \citet{reardon_portilla}, which found a decreasing gap for kindergarteners. This decline is evident also for 15 year olds, suggesting it might be more of an institutional change rather than a specific grade-level policy. Do note that the trendline is for both mathematics and literacy. If it were only for mathematics, then the trendline would probably be even more pronounced. Despite this, pooling both subjects gives even more strength to the analysis given that the results stand for both subjects together \footnote{Same graph is availabe for reading upon request, where we find pretty much the same results if not even more steep/downward slopes.}. We also plot the 80/20 and 70/30 SES gaps and find very similar patterns (see figure 13 in the appendix).

@@

We find that the initial gap for the U.S in 2000 is 2.17 varying between 2.13 and 2.21. This gap is much higher than found in previous studies, but again, we don't know how much previous estimates vary and nobody has really studied the 15-year old achievement gap in isolation. All country-level studies pool many studies together and control for age, removing the age-specific gap effect.

Figure 3 takes a more direct approach and looks at the percentage change from the first and last timepoint available. Each data point has been computed together with it's 50\% and 95\% uncertainty interval \footnote{Each of these uncertainty intervals were computed using a 500-replicate bootstrap}.

@@

The U.S shows a different pattern from all other countries. Both the 90th/50th/10th gaps seem to contribute rather similarly to the overall 90th/10th gap (although 50th/10th is marginally higher) but more interestingly, the 90th/50th gap has stagnated in the last 15 years. The top/middle groups have a relatively big gap and it hasn't changed at all. On the other hand, the steep decrease we saw for the U.S is coming exclusively from the 50th/10th gap shrinking. The reason, nonetheless, might be different from the German case, where it's probable that the bottom is catching up. Here, as in the 90th/10th gap, the middle might be 'dumbing' down closer to the low SES group. The surprising result is that the reductiong of U.S gap is coming from the shrinking of the 50/10 gap and not the 90/10 gap. These results are puzzling. Does this mean that the overall distribution of inequality in the U.S is being reduced toward the benefit of the Low SES group? The reduction is not uniform across all SES groups. We see that the 90/10 gap is closing, yet the the 90/50 gap is very stable over time. Despite this, the 50/10 is also shrinking, meaning that the 90 and 50 SES groups are shrinking down relative to the bottom while the bottom is coming up. This is something we can only speculate but the evidence points toward something like that.

@@

To sum up, we find no evidence in support of our first hypothesis. We uncover that the tracking setup of a country doesn't imminently mean a certain degree of equality/inequality. Countries where the top SES group is increasing and the bottom is decreasing (top-left panel) have dramatically different tracking setups. Austria is an early tracker with several tracks, while Sweden is a late tracker with only 1 track. On the opossite panel (bottom-right), where we find countries that the bottom group is catching up while the top is coming down, we find similar results. Denmark and Germany are opposite poles in terms of tracking, while the U.S is distinctive in that it has little formal curricular stratification. We see no definitive evidence of top-to-bottom equality in all of these countries (except for Poland). We find that regardless of the different tracking setups, there's a tradeoff: either the bottom comes up and the top comes down or the top goes even further up and the bottom further down.

@@

Moving on, we now discuss if there's a tradeoff between an educational system being efficient, i.e raising the overall performance of the country, and whether that challenge compromises equality at the same time. For example, in order to improve the average score in, let's say, mathematics, tracking can improve the average score by making children in the higher trackers perform better, raising the overall performance. This, however, can come at a price. If tracking actually benefits the top but doesn't increase the performance of the bottom tracks then we see a net increase of inequality accompanied by an increase in the overall performance. Given that we have the data for the average performance for each country and the achievement gap, we can investigate this relationship in detail. We first being by plotting all countries pooled for each year in figure 8.

@@

For the past 15 years we see a very strong correlation suggesting that countries that perform very well in PISA have actually very low achievement gaps (or the other way around) \footnote{We graph all countries-years pooled as well and the relationship is identical}. The correlations are between -0.73 and -0.88. This confirms the theoretical and empirical frameworks suggested by \citet{werfhorst_mijs}. However, we should not get trapped into an ecological fallacy and assume this relationship is identical within each country. With the available data we can look at the evolution separately for countries. We plot it up next in figure 9.

@@

In the final model, we regress the inequality of a country on a dummy variable where 1 equals only 1 track vs more than 1 track, whether the age of first selection is 15 or more vs below 15, a dummy stating whether the percentage of tracked curriculum is above 0 (that is, any tracking), and the standardized vocational index \footnote{We create dummies to avoid multicollinearity. The final model has a maximum VIF of 2.2.}. We are reluctant to include more variables in the model, first to keep preserve parsimony, and secondly to prevent overfitting due to our sample size. We present the results in table 2.

@@

Given that the classical p values are inexistent in bayesian modelling, we present uncertainty intervals next to each coefficient. The first model only includes the 1 track dummy. This variable shows that countries that have one track, have on average, half a standard deviation of less inequality than other countries. This covariate alone explains about 23\% of the variance of the evolution of the 90/10 achievement gap. The second model now includes the dummy of whether the country has an age selection of 15 or more and we can see that countries with late selection into tracking have about -.38 less standard deviations in inequality. Also note that the number of tracks dummy was now reduced, suggesting these two variables are explaining similar things. This is expected given that late tracking is usually associated with fewer tracks, such as in the Scandinavian countries. The third model now includes whether the country has any degree of tracking, meaning whether the country has above 0\% of the curriculum tracked. The coding of the variable comes from visual inspection of the distribution, where some countries, such as the United States and Canada, have untracked curriculums. For this last variable we see no relationship at all, but after including the vocational index, the model is completely reversed with countries with tracked curriculum having about 0.6 more standard deviations in inequality that countries with no tracking.

We also note that vocational enrollment is associated with less inequality. Not surprisingly, the age of selection coefficient also increased due to the correlation between age of selection and vocational enrollment. Countries with high vocational enrollment are also those with early age selection, such as Germany and Austria. These two variables should be inspected further, perhaps with an interaction, given that both coefficients are sizable and very correlated. This final model explains over one third of the variance in the evolution of the 90/10 SES gap, with an R square of 40\%. The uncertainty of each of the covariates varies a lot with the percentage of curriculum tracked dummy ranging all the way from 0.9 to 0.3. Nonetheless, we see that the other coefficients are very strong and have reasonable intervals. It is very important to consider the uncertainty of our estimates, to gain confidence in our model fit. So far, this model seems to fit the data quite reasonably.

All of these countries have a different tracking structure, with some setups being more egalitarian than others. Using the previous model we make a simulation and predict the level of the 90/10 achievement gap for all possible tracking setups for all countries. In figure 10 we plot all these combinations. The X axis has the number of tracks that a country has, the Y axis has the predicted 90/10 achievement gap and below each country name we have whether the country has an age selection of 15 or more or below 15. In addition, to confirm how accurate the model is, we also plot the actual level of inequality in the actual tracking setup of the country. Let's take Germany for example.

@@

These graph reveals some interesting patterns. Generally speaking, we find that the simulation predicts that virtually all countries which are not in this 'ideal' tracking, that would switch to the 'ideal' setup, would experience a reduction of the achievement gap. Yet we also find the opposite result. Countries which are in the 'ideal' setup that would change to the 'worst' setup would see a widening of the achievement gap. On average, we find that if all countries switched to the 'ideal' setup, we would see an average reduction of 20\%. This is a very important figure considering that \citet{reardon2011} found that the US gap increased by about 40-50\% in 30 years. Had all countries in the 'ideal' tracking switched to the 'worst' tracking, we would see an average widening of nearly 72\%. For those countries which are not in the 'ideal' tracking setup, we compute the percentage change from switching to the ideal tracking in table 3.

@@

A more summarized empirical strategy is to actually standardize all the tracking indicator into one index, something that \citet{bol2013} already do quite nicely. Using this standardized tracking variable and the same bayesian multilevel model, we now model the tracking and vocational index as before, but add an interaction between two, given that in the previous model we recorded a strong shift in the model when tracking-related variables and the vocational model where both included in the same model. Table 4 shows the results for this model.

@@

We confirm the results from the previous models, where more tracking is associated with an achievement gap of about 0.4 standard deviations wider and more vocational enrollment is associated with a reduction of about 0.2 standard deviations, half the slope of tracking. Could it be that the tracking effect might be offsetting the effect of the vocational index given that it's half the size of the slope?. Both these coefficients are actually significantly different from each other, so it is possible. Before we explore it visually, note that the interaction between the two is very reliable and significant, considering the small sample size. Figure 11 plots the interaction for different quantiles of the vocational index.

@@

{\centering \includegraphics[width=6.5in,height=5in]{figure/interaction_plot-1}

@@

The results are very interesting. We see that for lower levels of tracking (meaning little to no tracking) there's a very low achievement gap, regardless of whether the country has very high or very low vocational enrollment. However, once tracking enters the picture, vocational enrollment can be a strong equalizer. We see that the bottom quantile of vocational enrollment has almost twice the achievement gap as the top quantile for vocational enrollment. Moreover, our earlier interpretation (that tracking is more important than vocational enrollment for reducing the achievement gap) still holds and is very evident here. Despite the equalizing power of vocational enrollment, a high level of tracking with the highest level of vocational enrollment still leaves a country with an achievement gap of over 1.5 standard deviations. However, a country with low levels of tracking and \textit{any} level of vocational enrollment is nearly 1.3 standard deviations at best. A key finding from these results is that tracking is a reasonable explanation of the 90/10 achievement gap, and we should consider it as a very important component in the reduction of the achievement gap. However, these results highlight that it should not be studied in isolation from other features such as the degree of vocational enrollment. We see a very sizable change in the achievement gap for a given country with the same level of tracking and high and low levels of vocational enrollment.

@@

To test whether these results hold for other samples, we compute the same models as before but also for the 80/20 and 70/30 SES gap. Figure 12 plots these coefficients for the three SES gaps.

@@

Instead of estimating the evolution over time, we calculate the cumulative change over time for each country, summarizing the 6 year-time span of achievement gaps into one cumulative inequality gap. This approach dramatically reduces the total sample size to 32, the number of total countries available \footnote{We've been presenting a selected number of countries given the lack of space, but for every model/estimation we include all countries available.}. It is for this reason that we only include at most 2 covariates in the model, the tracking and vocational index, which both summarizes the previous models and keeps it parsimonious.

@@

\section{Conclusion}

The literature on achievement inequality has recently started studying the evolution of the achievement gap and it has unconvered a wide variety of results. The main finding is that the U.S gap has been widening over time by about 40-50\% from 1970 until the 2000. Others find that this recent trend has reversed, at least for an age-specific group since the 2000's. From an international perspective, recent studies have shown that there is wide variability in this trend, with some countries showing decreases while others increases. This paper looks to investigate the magnitud of these gaps across selected countries, and attempts to explain why they are happening.

We start off by looking at the evolution of the achievement gap and document the same variation for the 15-year old achievement gap across countries. The United States has been closing its 90/10 gap by about 50\% with a wide interval going down to nearly 35\%. This reduction is also present in Germany, quite shockingly a country with a very institutionalized tracking system. Other countries such as France and Austria experience the contrary with persistent increases in the achievement gap. Moreover, once we disaggregate these patterns into a more fine-graned analysis, we discover that the dynamics are very different. The U.S gap is shrinking because both the 90th and 50th SES groups are actually going down while the bottom group is coming up. These results are evident because the 90/10 and 50/10 gap are shrinking at a very rapid rate, whereas the 90/50 gap is flat in the same 15 year interval. Germany, which has a very similar reduction, portrays a different explanation. All three achievement gaps are shrinking suggesting a more uniform reduction. On the other hand, for countries which are experiencing a widening of the gap, we see that it's happening at all instances of the distribution; all SES groups show a very steep increase.

We then assess whether this gap is correlated with the country's average performance, and find a very strong correlation at the aggregate level (that is, all year-country combination pooled). Yet when we look at this relationship within each country, we find exactly the same pattern: countries with high performance show low inequality, whereas countries with low performance show high inequality. This gives greater support to the work of \citet{werfhorst_mijs}. Further research should attempt to disentangle the causality between these two opposite forces: is efficiency trumping inequality, or is it the other way around?

We complete the analysis by studying what explains the evolution of this inequality and we find that the degree of tracking of a country seems to be a possible explanation, but it shouldn't be studied without factoring in the degree of vocational enrollment. We find that if a country has a low level of tracking, vocational enrollment will not help decrease the achievement gap. Conversely, once tracking enters the picture, vocational enrollment can help unease the burden of inequality and reduce the achievement gap very strongly. In a similar vein, we use this model to peform a simulation for each country: if every country switched to little or no tracking, what would be the resulting reduction in the 90/10 achievement gap? We find that if all countries switched, we would find that on average countries would see their gap shrink by about 20\%, with countries like Hungary and Germany experiencing reductions of 46\% and 36\% respectively. We contrast the former simulation with another counterfactual scenario: what would be the average gap increase if all countries switched to a system where tracking predominates? We record an average widening of 72\%, something that suggests that increasing tracking is more than three times worse than reducing it.

We conclude by showing that tracking and vocational enrollment are playing an important role in explaining the evolution of the achievement gap. We find that these results are very suggestive that institutional features, more specifically, stratified curriculums, might be playing a very sizable role in the promotion of inequality. After having ran a battery of tests and different empirical strategies, the results are consistent and point out that we should pay attention to the source of the achievement gaps and study whether particular SES groups are contributing more to the gap than others. Aside from the results, this paper also revelead some puzzling questions that we leave up to future research to study.

@@

\begin{figure}[!htbp]
